COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-05-17 main.tex
Clean up resonance labels in figures,, finish outlining paper

Add justification for the dust circularizing here as well.

@@

Operating under the assumption that any dust produced in a collision will couple to the gas and circularize on a short timescale, we use the cylindrical distance at which a collision occurs to predict the radial structure of the dust profile. This is shown in figure \ref{fig:coll_hist_r}, which was constructed in a similar way to figure \ref{fig:coll_hist_a}. Most strikingly, the bump that was present near the 2:1 resonance is no longer present in cylindrical distance space, although it suddenly appears again in the $e_{hi}$ simulation. The dip that is present near the 3:1 MMR also disappears and instead presents as a bump in the $e_{hi}$ case. To determine how the resonant bodies actually contribute to the radial collision profile, we excluded collisions that fall between $2.495 < a < 2.505$ AU (near the 3:1 resonance) and between $3.2 < a < 3.35$ AU (near the 2:1 resonance), which is shown by the orange curve. Qualitatively, none of the bump or dip features present are changed by making this exclusion. This suggests that the features near resonance seen in semimajor axis space become 'smeared out' in cylindrical distance space due to the large spread in eccentricity and orbital orientation.

@@

Although the nonaxisymmetric pileup of bodies near the edges of resonances has been seen previously \citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As mentioned previously, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following the pendulum analogy, this is equivalent to reaching the point where the pendulum becomes suspended in the vertical position. When the critical angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction \citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonance to be stable. Inside of resonance, the critical angles librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift direction changing sign across the resonance. The net result of this is that bodies near the edge of the resonance experience a torque from that planet that temporarily aligns their pericenters with the orbital phase of the planet. After the encounter ends, differential rotation of the disk slowly destroys the configuration over part of an orbit. This explains why the concentration of planetesimals appears to 'follow' the the planet.

In a disk with no secular forcing, this dynamical phenomena would produce a 'dip' near the center of the resonance and a 'bump' at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance (indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii. If the the aphelion distance of the inner edge and the perihelion distance of the outer edge cross, we should expect to see the central dip feature in the collision profile disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We attribute this to the longer synodic period, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:3 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would require much higher resolution observations. For these reasons, we focus on the 2:1 MMR as the main diagnostic indicator.

@@

\subsection{Simulated ALMA Observations}

Aaron's stuff goes here. Can the bumps and dips near the 2:1 MMR actually be resolved? What does the 3:1 resonance look like? Although we might not be able to use the 3:1 MMR to constrain mass or eccentricity, at least two resonances must be identified to determine where the 2:1 MMR actually is.

\subsection{Implications of a Bump vs Dip at the 2:1 MMR}

@@

\section{Discussion and Summary}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/bump_dip_diag.png}
    \caption{The presence of a dip or bump near the resonances can be used to constrain the mass and eccentricity
    of the perturbing planet. The color scale indicates the relative strength of the features, while the dashed lines
    indicate boundaries in parameter space where a dip or a bump will be produced at the resonance.\label{fig:bump_dip_diag}}
\end{center}
\end{figure*}

>>>>

@@

\subsection{Where Does the Dust End Up?}

To construct a radial dust profile from the collision statistics, we begin by making the assumption that any dust generated by collisions is strongly coupled to the gas. For a dust grain larger than the mean free path of the gas particles, the timescale to damp its eccentricity is given by \citep{1976PThPh..56.1756A}

\begin{equation}\label{eq:t_edamp}
    t_{e, damp} = \frac{8 s \rho}{3 e C_{D} \rho_{g} v_{th}}.
\end{equation}

\noindent where $m$, $e$ and $s$ are the mass, eccentricity and radius of the dust grain. $C_{D}$ is a drag coefficient which is of order unity, $\rho_{g}$ is the local density of the gas and $v_{th}$ is its thermal speed. At 3 AU, the gaseous component of the solar nebula has a density of $4 \times 10^{-10}$ g cm$^{-3}$ and a typical thermal speed of $10^{5}$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 1 mm dust grain with a density of 1 g cm$^{-3}$ and an eccentricity of 0.05, the damping timescale is $4 \times 10^{-3}$ years. This is orders of magnitude smaller than the collision timescale, therefore we conclude that collisionally generated dust grains will immediately couple to the gas.

Operating under this assumption, a map of the relative concentration of second-generation dust can be constructed from the cylindrical 
distance at which the collisions occur. This is shown in figure \ref{fig:coll_hist_r}. Similar to figure \ref{fig:coll_hist_a}, we use a KDE with a bandwidth of 0.02 AU to assemble the collision locations into a radial distribution. Most strikingly, the bump that was present near the 2:1 resonance is no longer present in cylindrical distance space, although it suddenly appears again in the $e_{hi}$ simulation. The dip that is present near the 3:1 MMR also disappears and instead presents as a bump in the $e_{hi}$ case. To determine how the resonant bodies actually contribute to the radial collision profile, we excluded collisions that fall between $2.495 < a < 2.505$ AU (near the 3:1 resonance) and between $3.2 < a < 3.35$ AU (near the 2:1 resonance), which is shown by the orange curve. Qualitatively, none of the bump or dip features present are changed by making this exclusion. This suggests that the features near resonance seen in semimajor axis space become 'smeared out' in cylindrical distance space due to the large spread in eccentricity and orbital orientation.

@@

Although the nonaxisymmetric pileup of bodies near the edges of resonances has been seen previously \citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As mentioned previously, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following the pendulum analogy, this is equivalent to reaching the point where the pendulum becomes suspended in the vertical position. When the critical angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction \citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonances to be stable. Inside of resonance, the critical angles librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift direction changing sign across the resonance. The net result of this is that bodies near the edge of the resonance experience a torque from that planet that temporarily drives their pericenters toward the current orbital position of the planet. After the encounter ends, differential rotation of the disk slowly destroys the configuration over part of an orbit. This explains why the concentration of planetesimals appears to 'follow' the the planet.

\citet{2016ApJ...818..159T} provides a different explanation for this phenomenon. Instead, the nonaxisymmetric structure is claimed to be a product of the path that a resonant test particle takes in a frame co-rotating with the planet (see figure 8.4 of \citet{2000ssd..book.....M}). This explanation, however, does not hold for a large collection of planetesimals, especially when there is no forced eccentricity. In such a case, the orbits of planetesimals near resonance would be randomly aligned, and the axisymmetric structure should become washed out. In the case with a perturbing planet planet on a circular orbit, this structure is still present (see figure 3 of \citet{2016ApJ...818..159T}).

With no secular forcing, this dynamical phenomena would produce a 'dip' in the radial surface density profile near the center of the resonance and a 'bump' at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance (indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii. If the the aphelion distance of the inner edge and the perihelion distance of the outer edge cross, we should expect to see the central dip feature in the collision profile disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We attribute this to the longer synodic period, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:2 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would require much higher resolution observations. For these reasons, we focus on the 2:1 MMR as the main diagnostic indicator.

\subsection{Observability of the Dust}

Orange curves in figure \ref{fig:coll_polar_e} show the simulated ALMA flux profiles generated with {\sc CASA} \citep{2007ASPC..376..127M}. Have Aaron add in some more text here describing exactly how the profiles were created.

@@

Can the bumps and dips near the 2:1 MMR actually be resolved? What does the 3:1 resonance look like? Although we might not be able to use the 3:1 MMR to constrain mass or eccentricity, at least two resonances must be identified to determine where the 2:1 MMR actually is.

@@

\section{Conclusions and Summary}

Gap features take on a different structure for exterior resonances \citet{2016ApJ...818..159T}, which might make this bump vs dip diagnostic only useful for the interior 2:1 MMR.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/bump_dip_diag.png}
    \caption{The presence of a dip or bump near the resonances can be used to constrain the mass and eccentricity
    of the perturbing planet. The color scale indicates the relative strength of the features, while the dashed lines
    indicate boundaries in parameter space where a dip or a bump will be produced at the resonance.\label{fig:bump_dip_diag}}
\end{center}
\end{figure}
